---
title: "correlation"
output: html_notebook
---


```{r}
# read all data
library(tidyverse)
eye.df = read_csv(".\\eye_processed\\eye.output.csv")  %>% mutate(id = as.numeric(id))

str(eye.df)
driving.df = read_csv(".\\RT.csv")
str(driving.df)
scl.df = read_csv(".\\scaled_scl.csv")%>% mutate(task = str_remove(task, "\\d")) 
str(scl.df)
hrraw.df = read_csv(".\\hr.csv") %>% filter(!grepl('0', task))%>% mutate(task = str_remove(task, "\\d")) %>% filter(task != "B")
pose.df = read_csv(".\\pose_mean.csv")
str(pose.df)
experience.df = read_csv(".\\Demographic.csv") %>% select(id, age, av_experience)
str(experience.df)
big.maneuver = read_csv(".\\all_feature_clasification.csv") %>% select(id, task, big_maneuver)

```

```{r}
#clean data 
library(janitor)
#clean data frame 
hr.df = hrraw.df %>% clean_names(., "snake") %>% select(id, task, mean_hr_beats_min, max_hr_beats_min, mean_rr_ms, rmssd_ms, sd_hr_beats_min, sdnn_ms, rr_tri_index, lf_ms_2, hf_ms_2, lf_hf_ratio, total_power_ms_2)

#select driving 
rt.df = driving.df %>% select(id, task, RT_adoff, RT_steer)
```



```{r}
#join dataframes
df = hr.df %>% left_join(eye.df, by = c("id", "task")) %>% left_join(scl.df, by = c("id", "task")) %>% left_join(rt.df, by = c("id", "task")) %>% left_join(pose.df, by = c("id", "task")) %>% left_join(experience.df, by = c("id"))
str(df)
```
```{r}
RT_adoff.df = df %>% select(-c(id, task, RT_steer, y_first_gaze, y_first_fix)) %>%  select(RT_adoff, everything())
RT_adoff.df_0.25 = RT_adoff.df %>% select(RT_adoff, drv_percentage, pupil_z, pupil_change, switch_count, entropy_log2, drv_scanpath_len, drv_convex_hull, drv_fix_count, drv_fix_dur, drv_grid_transition, drv_entropy_log2, drv_spatial_density, drv_NNI)

RT_steer.df = df %>% select(-c(id, task, y_first_gaze, y_first_fix,RT_adoff)) %>%  select(RT_steer, everything())
#RT_steer.df_0.25 = RT_steer.df %>% select(RT_steer, drv_percentage, ppupil_z,pupil_change,  fix_count, switch_count, grid_transition, entropy_log2)

library(Hmisc)
adoff_matrix = rcorr(as.matrix(RT_adoff.df), type = "spearman")
#adoff_matrix = rcorr(as.matrix(RT_adoff.df_0.25), type = "spearman")
adoff_corr.df = as.data.frame(adoff_matrix$r)%>% select(RT_adoff) %>% tibble::rownames_to_column(var = "metric")
#print(adoff_matrix)

steer_matrix = rcorr(as.matrix(RT_steer.df), type = "spearman")
#print(steer_matrix )
steer_corr.df=as.data.frame(steer_matrix$r) %>% select(1) %>% tibble::rownames_to_column(var = "metric") 

first_fix.df = df %>% select(-c(id, task,RT_adoff, RT_steer, y_first_gaze)) %>%  select(y_first_fix, everything())
first_fix_matrix = rcorr(as.matrix(first_fix.df), type = "spearman")
first_fix_corr.df=as.data.frame(first_fix_matrix$r) %>% select(1) %>% tibble::rownames_to_column(var = "metric")

first_gaze.df = df %>% select(-c(id, task,RT_adoff, RT_steer, y_first_fix)) %>%  select(y_first_gaze, everything())
first_gaze_matrix = rcorr(as.matrix(first_gaze.df), type = "spearman")
first_gaze_corr.df=as.data.frame(first_gaze_matrix$r) %>% select(1) %>% tibble::rownames_to_column(var = "metric")

corr.df = adoff_corr.df %>% left_join(steer_corr.df, by = c("metric")) %>% left_join(first_fix_corr.df, by = c("metric")) %>% left_join(first_gaze_corr.df, by = c("metric")) %>% drop_na() 

corr.df_0.25 = corr.df %>%   rowwise() %>% mutate(max_value = max(abs(c_across(-1)), na.rm = TRUE)) %>% filter(max_value >=0.25)


```


```{r}
#plot correlation 
library(corrplot)

png(filename = "RT_adoff_corr.png", width = 2400, height = 1600)
#corrplot(adoff_matrix$r, method = 'ellipse', order = 'AOE', type = 'upper')
corrplot.mixed(adoff_matrix$r, order = 'AOE')
dev.off()

png(filename = "RT_steer_corr.png", width = 2400, height = 1600)
corrplot.mixed(steer_matrix$r, order = 'AOE')
dev.off()

png(filename = "firstfix_corr.png", width = 2400, height = 1600)
corrplot.mixed(first_fix_matrix$r, order = 'AOE')
dev.off()


```

```{r}
#cosolidated dataset

write.csv(df,file = "all_feature_nosubjective.csv", row.names = FALSE)

```


```{r}
#logistic regression for correlation against big maneuver
df.corr = df %>% left_join(big.maneuver, by = c("id","task")) %>% select(-c(id, task, RT_steer,RT_adoff, y_first_gaze, y_first_fix)) %>%  select(big_maneuver, everything())

# Load the necessary libraries
library(caret)
library(MASS)



# Fit a full logistic regression model with all predictors
fullModel <- glm(big_maneuver ~ ., data = df.corr, family = binomial)

# Summary of the full model to check initial coefficients
summary(fullModel)

# Stepwise feature selection using both directions (forward and backward)
stepModel <- stepAIC(fullModel, direction = "both", trace = FALSE)

# Summary of the stepwise model to check selected features and their coefficients
summary(stepModel)
```

```{r}
#Linear mixed model for feature selection 
library(dplyr)
library(lme4)
library(lmerTest)
df.corr = df %>% dplyr::select(-c(task, RT_steer, y_first_gaze, y_first_fix)) %>% na.omit()

# Standardizing predictors (except the categorical ones like id and task)
df.corr_standard <- df.corr %>%
  mutate_at(vars(-id, RT_adoff), scale)


# Prepare the formula
fixed_effects <- names(df.corr)[!(names(df.corr) %in% c("id", "RT_adoff"))]
formula_str <- paste("RT_adoff ~", paste(fixed_effects, collapse = " + "), "+ (1 | id)")
formula <- as.formula(formula_str)
print(formula
      )
#calculate initial model
initial_model <- lmer(formula , data = df.corr)
summary(initial_model)

initial_model_scaled <- lmer(formula , data = df.corr_standard, na.action = na.omit)
summary(initial_model_scaled)

step_fm <- step(initial_model)
step_fm # Display elimination results
final_fm <- get_model(step_fm)

final.model = lmer(RT_adoff ~ mean_hr_beats_min + mean_rr_ms + pupil_z + entropy_log2 + spatial_density + (1 | id), 
                 data = df.corr)
summary(final.model)


```

```{r}

#logistic regression for correlation against big maneuver
df.corr = df%>% dplyr::select(-c(id, task, RT_steer,y_first_gaze, y_first_fix)) 
df.corr <- na.omit(df.corr)

#sum(is.na(df.corr))


# Load the necessary libraries
library(MASS)

# Fit a full logistic regression model with all predictors
fullModel <- lm(RT_adoff ~ ., data = df.corr)

# Summary of the full model to check initial coefficients
summary(fullModel)

# Stepwise feature selection using both directions (forward and backward)
stepModel <- stepAIC(fullModel, direction = "both", trace = FALSE)

# Summary of the stepwise model to check selected features and their coefficients
summary(stepModel)
```

```{r}
#logistic regression for correlation against big maneuver
df.corr = df%>% dplyr::select(-c(id, task, RT_adoff,y_first_gaze, y_first_fix)) 
df.corr <- na.omit(df.corr)

#sum(is.na(df.corr))


# Load the necessary libraries
library(MASS)

# Fit a full logistic regression model with all predictors
fullModel <- lm(RT_steer ~ ., data = df.corr)

# Summary of the full model to check initial coefficients
summary(fullModel)

# Stepwise feature selection using both directions (forward and backward)
stepModel <- stepAIC(fullModel, direction = "both", trace = FALSE)

# Summary of the stepwise model to check selected features and their coefficients
summary(stepModel)
```


```{r}

#logistic regression for correlation against big maneuver
df.corr = df%>% dplyr::select(-c(id, task, RT_steer,RT_adoff, y_first_gaze)) 
df.corr <- na.omit(df.corr)

#sum(is.na(df.corr))


# Load the necessary libraries
library(MASS)

# Fit a full logistic regression model with all predictors
fullModel <- lm(y_first_fix ~ ., data = df.corr)

# Summary of the full model to check initial coefficients
summary(fullModel)

# Stepwise feature selection using both directions (forward and backward)
stepModel <- stepAIC(fullModel, direction = "both", trace = FALSE)

# Summary of the stepwise model to check selected features and their coefficients
summary(stepModel)
```
```{r}

#logistic regression for correlation against big maneuver
df.corr = df%>% dplyr::select(-c(id, task, RT_steer,RT_adoff, y_first_fix)) 
df.corr <- na.omit(df.corr)

#sum(is.na(df.corr))


# Load the necessary libraries
library(MASS)

# Fit a full logistic regression model with all predictors
fullModel <- lm(y_first_gaze ~ ., data = df.corr)

# Summary of the full model to check initial coefficients
summary(fullModel)

# Stepwise feature selection using both directions (forward and backward)
stepModel <- stepAIC(fullModel, direction = "both", trace = FALSE)

# Summary of the stepwise model to check selected features and their coefficients
summary(stepModel)
```
Generalized Additive Models (GAMs)
While GAMs can be considered a bit more advanced than basic statistical models, they are highly interpretable because they model the relationship between predictors and the response variable using smooth functions. Each function's shape can give you insights into the nature of the relationship.

Pros:

Flexible enough to capture non-linear relationships.
Each predictor's effect is modeled separately, enhancing interpretability.
Cons:

Requires selecting an appropriate smoothing parameter, though methods like Generalized Cross Validation (GCV) can assist with this.
Slightly more complex than linear models, but still within the realm of basic statistics.

```{r}
library(gam)
library(broom)
library(tidyverse)

# Assume your data is in a dataframe called df
# Ensure 'experience' is a factor
gam_df <- df %>% select(-c(id, task, RT_steer, y_first_gaze, y_first_fix, av_experience)) 

# Create a formula for the gam model
# Automatically create the formula string with smooth terms for all variables except 'y' and 'experience'
predictors <- setdiff(names(gam_df), c("RT_adoff"))
formula_gam <- as.formula(
  paste("RT_adoff ~", paste("s(", predictors, ")", collapse=" + "))
)



# Fit the initial model
initial_gam <- gam(formula_gam, 
                     data = gam_df, 
                     family = gaussian())

summary(initial_gam)
# # Get a tidy summary of the model
# tidy_initial_gam <- tidy(initial_gam)
# print(tidy_initial_gam)
# 
# # Model selection with step.gam (Note: step.gam is not a tidyverse function, but we use it within a tidyverse workflow)
# mgcv::step_gam <- step.gam(initial_gam, scope = list(upper = formula_gam, lower = y ~ 1), direction = "backward")
# 
# # Get a tidy summary of the selected model
# tidy_step_gam <- tidy(step_gam)
# print(tidy_step_gam)
# 
# # Plot the effects of the selected smooth terms
# plot(step_gam) # This is not a tidyverse plot, but integrates well with the workflow
# 
# # Model diagnostics
# gam.check(step_gam) # This is not a tidyverse function, but is necessary for GAM diagnostics



```

