---
title: "driving_position"
output: html_notebook
---

This file focuses on the VDS_chassis_CG_position variable, using particular coordinates of the objects in the scerario to derive location, min_distance etc. 
the location of the first cone is (-5393.49, 7644.40)

```{r,message=FALSE}
#read in position files
library(tidyverse)
library(readr)
# directory with CSV files

#use relative path
root_dir <- ".\\driving\\driving_position\\"

# list all CSV files
all_csv_files <- list.files(path = root_dir, pattern = "\\.csv$", full.names = TRUE, recursive = TRUE)

# read all CSV files into a list of dataframes
list_df <- lapply(all_csv_files, function(x) {
  df <- read_csv(x)
  df$filename <- basename(x)   # file name
 # df$B <- dirname(x)    # directory name
  df
})

# bind all dataframes into one
driving_position.raw <- bind_rows(list_df)
```


## Clean and label data

```{r, message=FALSE}
#str(driving_position.raw)
driving_position.label <- driving_position.raw %>%
  mutate(
    #extract the task name from file name
    B = str_sub(filename, start = 10, end = 14),
    # extract task letter from file name
    task = case_when(
      str_starts(B, "Pu") ~ "P",
      str_starts(B, "Po") ~ "M",
      str_starts(B, "D") ~ "D",
      str_starts(B, "J") ~ "J",
      str_starts(B, "W") ~ "W",
      TRUE ~ B),
    #extract participant number, remove the letter P 
    id = str_replace(subject, pattern = "^P", replacement = ""), 
    #correct column type
    task = as.factor(task),
    id = as.numeric(id)
  ) %>% select(id, task, Frame, CG_x, CG_y)
str(driving_position.label)
```

```{r}
#check = driving_position.label %>% filter(is.na(CG_y))
#get TOR1 frame number 
library(tidyverse)
key_moments = read_csv(".\\RT.csv") %>% select(id, task, TOR1)

#define cone location (reference location after adjusting for the car dimension)
fixed_point_x <- -5398.85
fixed_point_y <- 7647.40 # near the center of the ego lane


#calculate the dataframe that contains the frame where the cross of borderline took place
distance_cone = driving_position.label %>% 
  left_join(key_moments, by = c("id", "task")) %>% 
  mutate(frame_adj = Frame-TOR1) %>%
  filter(frame_adj>=0 & frame_adj<900) %>% 
  group_by(id, task) %>% 
  #calculate distance to first cone 
  mutate(distance = sqrt((CG_x - fixed_point_x)^2 + (CG_y - fixed_point_y)^2)) %>% 
  # x is on the roadway direction, we now calculate the x distance between vehicle and cone when the vehicle center line is 3 feet to the left of the cone, suggesting an avoidance
  arrange(frame_adj) %>%  # Ensures that we consider the points in the order of increasing y values (to right side is increase)
  filter(CG_y <= (fixed_point_y-2)) %>%  # Keep only the points where y is below the threshold (half car wide to the left, being able to avoid collision)
  slice(1) %>% 
  # Keep only the first point that meets the condition
  mutate(x_distance =  fixed_point_x - CG_x ) %>% 
  # in cases where the driver hit the cone, set value to 0 
  mutate(min_headway = ifelse(x_distance < 0, 0, x_distance))


#clean up the vehicle trajectory 
trajectory.df = driving_position.label %>% 
  left_join(key_moments, by = c("id", "task")) %>% 
  mutate(frame_adj = Frame-TOR1) %>%
  filter(frame_adj>=0 & frame_adj<900) %>% 
  group_by(id, task) %>% 
  mutate(y_adj = -(CG_y - fixed_point_y),
         x_adj = CG_x - fixed_point_x)
  # 

```

#Now extract the vehcicle speed at the time the cross took place to calculate TTC
```{r}
library(tidyverse)
#first read in the raw driving data (not directly available in repo)
veh_speed.df = read_csv(".\\driving\\year3_driving_all.csv") %>% select(id, task, Frame, veh_speed)%>% mutate(id = as.numeric(id))


TTC.df = distance_cone %>% mutate(id = as.numeric(id)) %>% select(id, task, Frame, min_headway) %>% left_join(veh_speed.df, by = c("id", "task", "Frame")) %>% mutate(min_TTC = min_headway/(veh_speed * (5280/3600)))
```



# Now plot trajectory data
```{r}
library(plotly)
library(RSelenium)#for higher resolution
#draw position trajectory 
# Consolidating lines and computing confidence interval
consolidated_data <- trajectory.df %>%
  group_by(task, frame_adj) %>%
  dplyr::summarise(mean_x = mean(x_adj),
                   mean_y = mean(y_adj),
                   se_x = sd(x_adj)/sqrt(n()),
                   se_y = sd(y_adj)/sqrt(n()))

# Plotting consolidated lines with confidence intervals
consolidated_plot <- ggplot(consolidated_data, aes(x = mean_x, y = mean_y, color = task, fill = task)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_y-se_y, ymax = mean_y+se_y), alpha = 0.3) +
  labs(title = "Trajectory by Task", color = "Task") +
  #draw the position of the cone 
  theme_minimal()

# Displaying the consolidated plot
ggplotly(consolidated_plot)

ggsave(path = ".\\image",filename = "trajectory_mean.svg", width = 12, height = 6, device='svg', dpi=400)
```

```{r}
# creating individual plot 
df = trajectory.df%>% dplyr::filter(task == "P")

# Calculate mean offset for each frame
mean_offsets <- consolidated_data %>% dplyr::filter(task == "P")



P<- ggplot(df, aes(x = x_adj, y = y_adj, color = as.factor(id))) +
    geom_line(alpha = 0.5) +
    labs(title = "Driving Trajectory for Puzzle Condition") +
    geom_line(data = mean_offsets, aes(x = mean_x, y = mean_y), color = "#0575f6", size = 2) +
    ylim(-6,24)+
    theme_minimal() 
  


ggplotly(P)
ggsave(path = ".\\image",filename = "trajectory_puzzle.svg", width = 12, height = 6, device='svg', dpi=400)

#########################################

# creating individual plot 
df = trajectory.df%>% dplyr::filter(task == "D")

# Calculate mean offset for each frame
mean_offsets <- consolidated_data %>% dplyr::filter(task == "D")



D<- ggplot(df, aes(x = x_adj, y = y_adj, color = as.factor(id))) +
    geom_line(alpha = 0.5) +
    labs(title = "Driving Trajectory for Drive Condition") +
    geom_line(data = mean_offsets, aes(x = mean_x, y = mean_y), color = "#f8786e", size = 2) +
    ylim(-6,24)+
    theme_minimal() 
  


ggplotly(D)
ggsave(path = ".\\image",filename = "trajectory_drive.svg", width = 12, height = 6, device='svg', dpi=400)

#########################################

# creating individual plot 
df = trajectory.df%>% dplyr::filter(task == "J")

# Calculate mean offset for each frame
mean_offsets <- consolidated_data %>% dplyr::filter(task == "J")



J<- ggplot(df, aes(x = x_adj, y = y_adj, color = as.factor(id))) +
    geom_line(alpha = 0.5) +
    labs(title = "Driving Trajectory for Jigsaw Condition") +
    geom_line(data = mean_offsets, aes(x = mean_x, y = mean_y), color = "#a7a90b", size = 2) +
    ylim(-6,24)+
    theme_minimal() 
  


ggplotly(J)
ggsave(path = ".\\image",filename = "trajectory_jigsaw.svg", width = 12, height = 6, device='svg', dpi=400)

#########################################
# creating individual plot 
df = trajectory.df%>% dplyr::filter(task == "W")

# Calculate mean offset for each frame
mean_offsets <- consolidated_data %>% dplyr::filter(task == "W")



W<- ggplot(df, aes(x = x_adj, y = y_adj, color = as.factor(id))) +
    geom_line(alpha = 0.5) +
    labs(title = "Driving Trajectory for Letter Condition") +
    geom_line(data = mean_offsets, aes(x = mean_x, y = mean_y), color = "#e76ff3", size = 2) +
    ylim(-6,24)+
    theme_minimal() 

ggplotly(W)
ggsave(path = ".\\image",filename = "trajectory_letter.svg", width = 12, height = 6, device='svg', dpi=400) 
##########################################

# creating individual plot 
df = trajectory.df%>% dplyr::filter(task == "M")

# Calculate mean offset for each frame
mean_offsets <- consolidated_data %>% dplyr::filter(task == "M")



M<- ggplot(df, aes(x = x_adj, y = y_adj, color = as.factor(id))) +
    geom_line(alpha = 0.5) +
    labs(title = "Driving Trajectory for Poem Condition") +
    geom_line(data = mean_offsets, aes(x = mean_x, y = mean_y), color = "#00bf7d", size = 2) +
    ylim(-6,24)+
    theme_minimal() 

ggplotly(M)
ggsave(path = ".\\image",filename = "trajectory_poem.svg", width = 12, height = 6, device='svg', dpi=400) 


```


```{r}
#use TTC.df as the output df, as it contains everything needed.
output.df <- TTC.df
```


```{r}
write_csv(output.df, "driving_position.csv")
```


## plot distance overtime 
```{r}
library(dplyr)
library(ggplot2)
library(plotly)
# Consolidating lines and computing confidence interval
consolidated_data <- distance_cone %>%
  group_by(task, frame_adj) %>%
  dplyr::summarise(mean_distance = mean(distance_x),
            lower_ci = mean(distance_x) - sd(distance_x)/sqrt(n()),
            upper_ci = mean(distance_x) + sd(distance_x)/sqrt(n()))

# Plotting consolidated lines with confidence intervals
consolidated_plot <- ggplot(consolidated_data, aes(x = frame_adj, y = mean_distance, color = task, fill = task)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.3) +
  labs(title = "distance to obstacle by Task", color = "Task") +
  theme_minimal()

# Displaying the consolidated plot
ggplotly(consolidated_plot)
```





Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
